Capstone Movielens Report

### 2025-12-11 reports/capstone-movielens.site/02-3.OMR-model.Rmd

  - [naive_model_MSEs](#func.naive_model_MSEs) (described in section [OMR Model: Helper Functions]) to compute *Mean Squared Errors* with the use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use `K = 5`);

>>>>

  - [naive_model_MSEs](#func.naive_model_MSEs) (described in section [OMR Model: Helper Functions]) to compute *Mean Squared Errors* with the use of *K-Fold Cross-Validation*, where the $K$ is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use $K = 5$);

### 2025-12-11 reports/capstone-movielens.site/02.5.UME-model.Rmd

Below, we provide the most significant part of the code for training our model with the use of *K-Fold Cross-Validation*, where the `K` is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use `K = 5`):
```{r eval=FALSE}
  cv.UM_effect <- train_user_movie_effect.cv()
```

>>>>

Below, we provide the most significant part of the code for training our model with the use of *K-Fold Cross-Validation*, where the $K$ is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use $K = 5$):
```{r eval=FALSE}
  cv.UM_effect <- train_user_movie_effect.cv()
```

### 2025-12-11 reports/capstone-movielens.site/02.6.UMGE-model.Rmd

Below, we provide the [line of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L902) for training our model using the `K-Fold Cross Validation` method, where the `K` is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use `K = 5`):
```{r, fn.train_user_movie_genre_effect.cv,   eval=FALSE}
cv.UMG_effect <- train_user_movie_genre_effect.cv()
```

>>>>

Below, we provide the [line of code](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L902) for training our model using the *K-Fold Cross Validation* method, where the $K$ is the length of the [`edx_CV` Object] (described in detail in [Appendix B: Models Training Datasets]) to which the *K-Fold Cross-Validation* is applied (in *this Project* we use $K = 5$):
```{r, fn.train_user_movie_genre_effect.cv,   eval=FALSE}
cv.UMG_effect <- train_user_movie_genre_effect.cv()
```

### 2025-12-11 reports/capstone-movielens.site/04.conclusion.Rmd

In this project, a _Machine Learning Algorithm_ was successfully developed for a _Recommender System_ (hereinafter referred to as _RS MLA_) to predict movie ratings using the [10M version of the MovieLens](http://grouplens.org/datasets/movielens/10m/) dataset.

>>>>

In this project, a _Machine Learning Algorithm_ was successfully developed for a _Recommender System_ (hereafter referred to as _RS MLA_) to predict movie ratings using the [10M version of the MovieLens](http://grouplens.org/datasets/movielens/10m/) dataset.

### 2025-12-11 reports/capstone-movielens.site/05.appendix.Rmd

Data frame representing the *UMGYDE Model* trained (and also *regularized* if the [lambda](#func.train_UMGY_SmoothedDay_effect.cv.args.lambda) argument value is non-default) with use of *K-Fold Cross-Validation*, where the $K$ is the length of the [`edx_CV` Object] (in *this Project* we use $K = 5$).
:::

@@

::: {.noteblock data-latex=""}
This helper function is used internally in the [UMGY_SmoothedDay_effect.predict.cv](#func.UMGY_SmoothedDay_effect.predict.cv) function (described below in this section) as shown in the [Usage](#func.UMGY_SmoothedDay_effect.predict.usage) subsection of *This Function's Description Section* below.
:::

@@

::: {.noteblock data-latex=""}
In the code snippet above, the function `UMGY_SmoothedDay_effect.predict` is called internally from the [UMGY_SmoothedDay_effect.predict.cv](#func.UMGY_SmoothedDay_effect.predict.cv) function described below in this section.
:::
:::

@@

:::
::: { #func.UMGY_SmoothedDay_effect.predict.args.UMGYD_effect .sidebar }
  - **UMGYD_effect:** Data frame representing the *UMGYDE Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYDE Model* passed in the [UMGYD_effect](#func.UMGY_SmoothedDay_effect.predict.args.UMGYD_effect) argument using the validation dataset passed in the [test_set](#func.UMGY_SmoothedDay_effect.predict.args.test_set) arg. 
:::

@@

@@

The source code of the [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L125) function is shown below:
```{r eval=FALSE}
```

>>>>

Dataframe representing the *UMGYDE Model* trained (and also *regularized* if the [lambda](#func.train_UMGY_SmoothedDay_effect.cv.args.lambda) argument value is non-default) with use of *K-Fold Cross-Validation*, where the $K$ is the length of the [`edx_CV` Object] (in *this Project* we use $K = 5$).
:::

@@

Predicts *user rating* values on the *Test dataset* passed in the [test_set](#func.UMGY_SmoothedDay_effect.predict.args.test_set) argument based on the data of the trained *UMGY Effect (UMGYDE) Model* passed in the [day_smoothed_effect](#func.UMGY_SmoothedDay_effect.predict.args.day_smoothed_effect) argument.

::: {.noteblock data-latex=""}
This function is also called within other helper functions, such as the [calc_UMGY_SmoothedDay_effect.MSE](#func.calc_UMGY_SmoothedDay_effect.MSE) and [UMGYDE_model.predict](#func.UMGYDE_model.predict) described below in this section and [Matrix Factorization Functions] section of *this Appendix*, respectively.
:::

@@

```{r eval=FALSE}
final.UMGYDE.predicted <- final_holdout_test |>
  UMGY_SmoothedDay_effect.predict(rglr.day_smoothed_effect)
```
:::

@@

:::
::: { #func.UMGY_SmoothedDay_effect.predict.args.day_smoothed_effect .sidebar }
  - **day_smoothed_effect:** Data frame representing the *UMGYDE Model*.

@@

This function implements an algorithm for computing the *user rating values* in accordance with the *mathematical description* of the *UMGYDE Model*, as described in section [Mathematical Description of the UMGYDE Model] of *this Report*.
:::

@@

Dataframe containing *predicted user rating values* based on the *UMGYDE Model*
:::

@@

The source code of the [UMGY_SmoothedDay_effect.predict](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L125) function is shown below:
```{r eval=FALSE}
UMGY_SmoothedDay_effect.predict <- function(test_set, day_smoothed_effect) {
  test_set |>
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(rglr.UMGY_effect, by='year') |>
    left_join(day_smoothed_effect, by='days') |>
    mutate(predicted = clamp(mu + a + b + g + 
                               ifelse(is.na(ye), 0, ye) +
                               ifelse(is.na(de_smoothed), 
                                      0, 
                                      de_smoothed))) |> 
    select(userId, movieId, timestamp, rating, predicted)
}
```
